---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---

## Loading and preprocessing the data
Read the data and unzip the file

```{r loading, message=FALSE, echo = TRUE, warning=FALSE}
# Read the data
data <- read.csv(unz("activity.zip", "activity.csv"), quote="\"", sep=",")
```

Required packages

```{r packages, message=FALSE, echo = TRUE, warning=FALSE }
library(ggplot2)
library(plyr)
library(dplyr)
```

Processing the data

```{r as.date, echo = TRUE }
data$as.date<- as.POSIXct(data$date, format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day

```{r histogram - plot.1, echo = TRUE, raw=true}
# Plot 1
hist(tapply(data$steps, data$as.date, sum), main = "Histogram total number of steps taken each day", xlab = "Steps")

```

The mean and median number of steps

```{r mean and madian, echo = TRUE }
steps_per_day <- aggregate(data$steps ~ data$date, FUN = sum)
colnames(steps_per_day)<- c("date", "steps")
mean(steps_per_day$steps)
median(steps_per_day$steps)
```
## What is the average daily activity pattern?
Time series plot of the average number of steps taken

```{r time series - plot.2, echo = TRUE,raw=true}

average <- ddply(data[!is.na(data$steps),], "interval", summarize, mean = mean(steps))
# Plot 2
ggplot(average, aes( x=interval, y=mean)) + geom_line() + xlab("Interval") + ylab("Average number of steps") + ggtitle("Average daily activity pattern")

```

The 5-minute interval that, on average, contains the maximum number of steps

```{r time series, echo = TRUE }
maximum <- filter(na.omit(average), mean == max(mean))
 maximum[1]
```

## Imputing missing values
Number of NA's value

```{r NA, echo = TRUE }
summary(data)
```

Strategy for imputing missing data: the mean of the 5-minute interval

```{r replace NA value, echo = TRUE }
# Calculate the mean of the 5-minute interval; replace NA value
average_2 <- ddply(na.omit(data), .(interval), summarize, steps = mean(steps))
na.data <- data[is.na(data$steps),]
data_2 <- merge(average_2, na.data, by=c("interval"))
data_2 = data_2 %>%
         select(steps.x, date, interval, as.date)
colnames(data_2)<- c("steps", "date", "interval", "as.date")
data_3 <- rbind(data[!is.na(data$steps),], data_2)
```

```{r histogram - plot.3, echo = TRUE, raw=true}
# Calculate the total number of steps taken per day
# Plot 3
par(mfrow=c(2,1))
hist(tapply(data$steps, data$as.date, sum), main = "Histogram total number of steps taken ", xlab = "Steps", ylim = c(0,50))
hist(tapply(data_3$steps, data_3$as.date, sum), main = "Histogram total number of steps taken- Imputed", xlab = "Steps", ylim = c(0,50), col = "blue")

```

## Are there differences in activity patterns between weekdays and weekends?

```{r convert date in weekday, echo = TRUE }
data_3$weekdays <- weekdays(as.Date(data$as.date))
data_3$datetype <- ifelse(data_3$weekdays %in% c("sábado", "domingo"), "Weekend", "Weekday")
 
average_2 <-subset(data_3, data_3$datetype == "Weekend")
average_2 <- ddply(average_2, "interval",  summarize, mean = mean(steps))
 
average_4 <-subset(data_3, data_3$datetype == "Weekday")
average_4 <- ddply(average_4, "interval",  summarize, mean = mean(steps))
```

```{r - plot 4, echo = TRUE, raw=true}
par(mfrow=c(1,1))
# Plot 4
plot(average_2$interval, average_2$mean, type = "l", ylab = "Average number of steps", xlab = "Interval", main = "Average daily activity pattern" , col = 2)
points(average_4$interval, average_4$mean, type = "l", col = 20)
legend("topright", c("Weekend", "Weekday"),lty=c(1,1), col= c(2,20))
```

